%% start of file `coverletter.tex'.
%% Copyright 2014 Francois Mouton (moutonf@gmail.com).
%
% This template is adapted from the work performed by Xavier Danaux (xdanaux@gmail.com).
%
% This work may be distributed and/or modified under the
% conditions of the LaTeX Project Public License version 1.3c,
% available at http://www.latex-project.org/lppl/.

\recipient{Department of Mathematics}{Pennsylvania State University}
\date{\today}
\opening{Dear Selection Committee,}
\closing{Sincerely,}
%\enclosure[Attached]{curriculum vit\ae{}}          % use an optional argument to use a string other than "Enclosure", or redefine \enclname
\makelettertitle

\justifying 

I am writing because I would like to be considered for your advertised position of postdoctoral scholar, to begin in the academic year 2020-21.

I am a PhD student at the Massachusetts Institute of Technology, graduating in Spring 2020, and my research is on dispersive partial differential equations and harmonic analysis. A project I am currently working on concerns the modeling of rogue ocean waves via the Dysthe equation. Despite the promising physics literature, even basic properties like local and global well-posedness for the two dimensional equation are unknown, among other interesting questions. Some of my past work has to do with the continuum limit of discrete NLS-type equations, as well as developing a local well-posedness theory for dispersive PDEs with a memory effect. Given our common research interests, I am very excited about the possibility to interact with the strong faculty at Penn State, especially Prof. Nguyen, among others.

During my years at MIT, I have had some teaching experience as a recitation instructor for multivariable calculus and for differential equations, I have held weekly problem sessions for a variety of other subjects, and I have mentored undergraduate students on intensive summer research projects. I have always received very positive teaching evaluations, and I am looking forward to further improving my teaching as I gain more experience.

A CV, a detailed research statement and a teaching statement are included with this application. Moreover, I have asked the following professors to submit a letter of recommendation:
\begin{list}{$\bullet$}{}
\item Luis Vega, Basque Center for Applied Mathematics.
\item David Jerison, Massachusetts Institute of Technology.
\item Gigliola Staffilani, Massachusetts Institute of Technology (thesis advisor). 
\item Jeremy Orloff, Massachusetts Institute of Technology (teaching reference).
\end{list}

Please do not hesitate to contact me should you have any questions. I look forward to hearing from you.

\hspace{1.5cm}

\makeletterclosing

\clearpage

%% end of file `coverletter.tex'.