%% start of file `coverletter.tex'.
%% Copyright 2014 Francois Mouton (moutonf@gmail.com).
%
% This template is adapted from the work performed by Xavier Danaux (xdanaux@gmail.com).
%
% This work may be distributed and/or modified under the
% conditions of the LaTeX Project Public License version 1.3c,
% available at http://www.latex-project.org/lppl/.

Example letter:
\recipient{Scientific and organizing committee}{Gran Sasso Quantum Meetings: from many particle systems to quantum fluids}
\date{\today}
\opening{Dear Sir or Madam,}
\closing{Yours faithfully,}
%\enclosure[Attached]{curriculum vit\ae{}}          % use an optional argument to use a string other than "Enclosure", or redefine \enclname
\makelettertitle

I am writing because I would like to participate in the upcoming Gran Sasso Quantum Meeting workshop taking place at the GSSI from November 28th to December 1st. I am a fourth-year graduate student at the Massachusetts Institute of Technology, doing my thesis on nonlinear dispersive PDEs under the supervision of Prof. Gigliola Staffilani, who let me know about this conference and encouraged me to apply.

There are many reasons why I think I would greatly benefit from this workshop.
First and foremost, I find the topics that will be presented very interesting. 
Furthermore, during my latest work I have realized that a deeper understanding of the physical and probabilistic viewpoint 
%of the equations that I am working with?
could help my research, which I am sure this workshop could provide, both during the courses and in conversations with other participants.
Finally, this workshop could give rise to possible future collaborations with some participants whose work I am very interested in, such as my advisor's previous student, Dr. Vedran Sohinger, or Prof. Juan J. L. Vel\'azquez.

I would also like to apply for some financial support for accommodation and local expenses, which would help to complement some funding for travel expenses offered by my own institution and advisor.

Please do not hesitate to contact me should you have any further questions.

\makeletterclosing

\clearpage

%% end of file `coverletter.tex'.